---
title: "Doubet detection with Scrublet"
author: 
- name: "Maria Lucia Romero Rivero"
  email: MLROMEROR@clinic.cat
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
    fig_captions: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message=FALSE, warning = FALSE)
options(width = 1200)
```

# Introduction

In this notebook, we summarize the doublet detection results obtained after running [Scrublet](https://www.sciencedirect.com/science/article/pii/S2405471218304745) for 
the multiome data.


# Histograms

We start by plotting the doublet score histograms for observed transcriptomes and simulated doublets.
We should visualize a bimodal distribution in the simulated doublets that allows to set an accurate threshold,
however we did gt any bimodal distribution.
/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_14_T_2/Figures

## BCLL_2_T_1(RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_2_T_1/Figures/scrublet_histogram.png"; sprintf("![](%s)", path_to_image)`

## BCLL_2_T_2(RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_2_T_2/Figures/scrublet_histogram.png"; sprintf("![](%s)", path_to_image)`

## BCLL_2_T_3(RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_2_T_3/Figures/scrublet_histogram.png"; sprintf("![](%s)", path_to_image)`

## BCLL_8_T_1 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_8_T_1/Figures/scrublet_histogram.png"; sprintf("![](%s)", path_to_image)`

## BCLL_8_T_2 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_8_T_2/Figures/scrublet_histogram.png"; sprintf("![](%s)", path_to_image)`

## BCLL_9_T_1 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_9_T_1/Figures/scrublet_histogram.png"; sprintf("![](%s)", path_to_image)`

## BCLL_9_T_2 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_9_T_2/Figures/scrublet_histogram.png"; sprintf("![](%s)", path_to_image)`

## BCLL_14_T_1 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_14_T_1/Figures/scrublet_histogram.png"; sprintf("![](%s)", path_to_image)`

## BCLL_14_T_2 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_14_T_2/Figures/scrublet_histogram.png"; sprintf("![](%s)", path_to_image)`

## BCLL_15_T_1 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_15_T_1/Figures/scrublet_histogram.png"; sprintf("![](%s)", path_to_image)`

## BCLL BCLL_15_T_2 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_15_T_2/Figures/scrublet_histogram.png"; sprintf("![](%s)", path_to_image)`


## BCLL-2-T (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/merged/matrix_with_BCLL_2/Figures/scrublet_histogram_peak.png"; sprintf("![](%s)", path_to_image)`


## non-BCLL-2-T (RNA)
`r path_to_image <-  "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/merged/matrix_without_BCLL_2/Figures/scrublet_histogram_peak.png"; sprintf("![](%s)", path_to_image)`


# UMAPs

As described in the best practices "Visualize the doublet predictions in a 2-D embedding (e.g., UMAP or t-SNE). Predicted doublets should mostly co-localize (possibly in multiple clusters). 

**If they do not, you may need to adjust the doublet score threshold, or change the pre-processing parameters to better resolve the cell states present in your data".**

## BCLL_2_T_1(RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_2_T_1/Figures/UMAP_doublets.png"; sprintf("![](%s)", path_to_image)`

## BCLL_2_T_2(RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_2_T_2/Figures/UMAP_doublets.png"; sprintf("![](%s)", path_to_image)`

## BCLL_2_T_3(RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_2_T_3/Figures/UMAP_doublets.png"; sprintf("![](%s)", path_to_image)`

## BCLL_8_T_1 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_8_T_1/Figures/UMAP_doublets.png"; sprintf("![](%s)", path_to_image)`

## BCLL_8_T_2 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_8_T_2/Figures/UMAP_doublets.png"; sprintf("![](%s)", path_to_image)`

## BCLL_9_T_1 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_9_T_1/Figures/UMAP_doublets.png"; sprintf("![](%s)", path_to_image)`

## BCLL_9_T_2 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_9_T_2/Figures/UMAP_doublets.png"; sprintf("![](%s)", path_to_image)`

## BCLL_14_T_1 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_14_T_1/Figures/UMAP_doublets.png"; sprintf("![](%s)", path_to_image)`

## BCLL_14_T_2 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_14_T_2/Figures/UMAP_doublets.png"; sprintf("![](%s)", path_to_image)`

## BCLL_15_T_1 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_15_T_1/Figures/UMAP_doublets.png"; sprintf("![](%s)", path_to_image)`

## BCLL_15_T_2 (RNA)
`r path_to_image <- "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/each_sample/BCLL_15_T_2/Figures/UMAP_doublets.png"; sprintf("![](%s)", path_to_image)`


## BCLL (RNA)
`r path_to_image <-  "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/merged/matrix_all/Figures/UMAP_doublets_peak.png"; sprintf("![](%s)", path_to_image)`


## BCLL-2-T (RNA)
`r path_to_image <-  "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/merged/matrix_with_BCLL_2/Figures/UMAP_doublets_peak.png"; sprintf("![](%s)", path_to_image)`


## non-BCLL-2-T (RNA)
`r path_to_image <-  "/Users/mlromeror/Documents/multiome_tonsil_Lucia/2.doublet_detection/tmp/merged/matrix_without_BCLL_2/Figures/UMAP_doublets_peak.png"; sprintf("![](%s)", path_to_image)`



# Session Information

```{r}
sessionInfo()
```

