---
title: "7- Filtering cluster 7"
author: 
- name: "Maria Lucia Romero Rivero"
  email: MLROMEROR@clinic.cat
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
    fig_captions: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pre-processing 

## Load packages

```{r message=FALSE, warning=FALSE}
library(Seurat)
library(Signac)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(magick)
library(knitr) 
library(kableExtra)
library(devtools)
library(harmony)
library(patchwork)
library(kableExtra)


```

## Parameters

```{r}
# Paths
path_to_obj <- ("~/Documents/multiome_tonsil_Lucia/results/R_objects/11.tonsil_multiome_integrated_without_doublets_normalized.rds")
path_to_save <- ("~/Documents/multiome_tonsil_Lucia/results/R_objects/12.tonsil_multiome_without_cluster7_doublets_normalized.rds")

```


## Load data

```{r eval=FALSE, include=FALSE}
tonsil_wnn <- readRDS(path_to_obj)
```

```{r fig.wide=TRUE}
vars <- str_subset(colnames(tonsil_wnn@meta.data), "^wsnn_res")
clusters_gg <- purrr::map(vars, function(x) {
  p <- DimPlot(
    tonsil_wnn,
    group.by = x,
    reduction = "wnn.umap",
    pt.size = 0.1, label = T
  )
  p 
})
clusters_gg
```


## Filtering  doublet cluster  7


```{r echo=TRUE}
tonsil_wnn$is_tcell <- 
   tonsil_wnn$wsnn_res.0.02 == "7"
  

tonsil_wnn <- subset(tonsil_wnn, subset = is_tcell == FALSE)


```

```{r echo=TRUE}
DimPlot(tonsil_wnn, reduction = "wnn.umap", group.by = "wsnn_res.0.02", label = TRUE, pt.size = 0.1)
```



## scATAC

### Normalization and linear dimensional reduction

 We exclude the first dimension as this is typically correlated with sequencing depth
Cells cluster completely separately in ATAC without harmony; so run harmony after SVD

RunSVD LSI
```{r}
DefaultAssay(tonsil_wnn) <- "peaks"
tonsil_wnn <- RunTFIDF(tonsil_wnn)
tonsil_wnn <- FindTopFeatures(tonsil_wnn, min.cutoff = "q0")
tonsil_wnn <- RunSVD(tonsil_wnn)
```



### Plot the Depth correlation plot

Compute the correlation between total counts and each reduced dimension component.

LSI component is typically highly correlated with sequencing depth. 
The first LSI component often captures sequencing depth (technical variation) rather 
than biological variation. If this is the case, the component should be removed from 
downstream analysis. We can assess the correlation between each LSI component and 
sequencing depth using the DepthCor() function:

For scRNA-seq data we don't typically observe such a strong relationship between the first PC and sequencing depth, and so usually retain the first PC in downstream analyses.


```{r fig.wide=TRUE}

DepthCor(tonsil_wnn)

```

Here we see there is a very strong correlation between the first LSI component and the total 
number of counts for the cell, so we will perform downstream steps without this component.


### UMAP representation

* dimensional reduction key, specifies the string before the number for the dimension names. UMAP by default
* reduction.name: Name to store dimensional reduction under in the Seurat object


```{r fig.wide=TRUE}
tonsil_wnn <- RunUMAP(
  tonsil_wnn,
  dims = 2:40,
  reduction = "lsi",
  reduction.name = "umap.atac",
  reduction.key = "atacUMAP_"
)
atac.umap<-DimPlot(
  tonsil_wnn,
  reduction = "umap.atac",
  group.by = "library_name",
  pt.size = 0.1
) + ggtitle('scATAC UMAP') + NoLegend()

atac.umap

#split_by: library ,edad, genero
```



## scRNA

### Normalization and linear dimensional reduction- 

### NormalizeData

```{r message=FALSE, warning=FALSE}
DefaultAssay(tonsil_wnn) <- "RNA"
tonsil_wnn <- NormalizeData(
  tonsil_wnn,
  normalization.method = "LogNormalize",
  scale.factor = 1e4
)

tonsil_wnn <- tonsil_wnn %>%
  FindVariableFeatures(nfeatures = 3000) %>%
  ScaleData() %>% 
  RunPCA() 
```



```{r}
PCAPlot(tonsil_wnn,
  group.by = "library_name")

```


```{r}
ElbowPlot(object = tonsil_wnn)
```


### UMAP representation

```{r}
tonsil_wnn <- RunUMAP(
  tonsil_wnn,
  dims = 1:30,
  reduction = "pca",
  reduction.name = "umap.rna",
  reduction.key = "rnaUMAP_"
)
rna.umap<-DimPlot(
  tonsil_wnn,
  reduction = "umap.rna",
  group.by = "library_name",
  pt.size = 0.1) + NoLegend() + ggtitle('scRNA UMAP')

rna.umap
```


```{r}
atac.umap + rna.umap
```


hacer primero harmony, quitar batch effect. 
atac y rna. harmony

# Run Harmony 

Pass the Seurat object to the RunHarmony function and specify which variable to integrate 
out. A Seurat object is returned with corrected Harmony coordinates.

## scATAC

```{r}
DefaultAssay(tonsil_wnn) <- "peaks"
tonsil_wnn <- RunHarmony(
  object = tonsil_wnn,
  reduction = "lsi",
  dims = 2:40,
  group.by.vars = "library_name",
  assay.use = "peaks",
  project.dim = FALSE,
  reduction.save = "harmony_peaks"
)
```


### UMAP representation

```{r}
tonsil_wnn <- RunUMAP(
  tonsil_wnn,
  dims = 2:40,
  reduction = "harmony_peaks",
  reduction.name = "umap.atac",
  reduction.key = "atacUMAP_"
)

Harm_peak<-DimPlot(
  tonsil_wnn,
  reduction = "umap.atac",
  group.by = "library_name",
  pt.size = 0.1
) + NoLegend() + ggtitle('Peak Harmony')
```

# scRNA

```{r message=FALSE, warning=FALSE, include=FALSE}
DefaultAssay(tonsil_wnn) <- "RNA"
tonsil_wnn <- RunHarmony(
  object = tonsil_wnn,
  reduction = "pca",
  dims = 1:30,
  group.by.vars = "library_name",
  assay.use = "RNA",
  project.dim = FALSE,
  reduction.save = "harmony_rna"
)
```

```{r}
tonsil_wnn <- RunUMAP(
  tonsil_wnn,
  dims = 2:40,
  reduction = "harmony_rna",
  reduction.name = "umap.rna",
  reduction.key = "rnaUMAP_"
)
Harm_rna<-DimPlot(
  tonsil_wnn,
  reduction = "umap.rna",
  group.by = "library_name",
  pt.size = 0.1
) + NoLegend() + ggtitle('RNA Harmony')
```

## scATAC and RNAseq Harmony

```{r}
Harm_peak+Harm_rna 
```


## Joint UMAP visualization

**FindNeighbors**

Constructs a Shared Nearest Neighbor (SNN) Graph for a given dataset. We first 
determine the k-nearest neighbors of each cell. We use this knn graph to construct 
the SNN graph by calculating the neighborhood overlap (Jaccard index) between 
every cell and its k.param nearest neighbors.

```{r}
# build a joint neighbor graph using both assays
tonsil_wnn <- FindMultiModalNeighbors(
  object = tonsil_wnn,
  reduction.list = list("harmony_peaks", "harmony_rna"),
  dims.list = list(2:40, 1:30), modality.weight.name = "Joint_snn_umap"
  )

# build a joint UMAP visualization

tonsil_wnn <- RunUMAP(
  object = tonsil_wnn,
  nn.name = "weighted.nn",
  reduction.name = "wnn.umap",
  reduction.key = "wnnUMAP_")
```

```{r}
joint.umap<- DimPlot(tonsil_wnn, label = FALSE, group.by = "library_name", pt.size = 0.1,  reduction = "wnn.umap") + plot_annotation(title = 'Joint UMAP')+ ggtitle('Joint UMAP by library name') + NoLegend()

joint.umap
```


```{r}
joint.umap_age<- DimPlot(tonsil_wnn, label = FALSE, split.by = "age_group", pt.size = 0.1,  reduction = "wnn.umap") + plot_annotation(title = 'Joint UMAP ')+ ggtitle('Joint UMAP age ') 

joint.umap_hospital<- DimPlot(tonsil_wnn, label = FALSE, split.by = "hospital", pt.size = 0.1,  reduction = "wnn.umap") + plot_annotation(title = 'Joint UMAP ')+ ggtitle('Joint UMAP age ') 

joint.umap_age
joint.umap_hospital
```


## FindClusters

```{r}
#find cluster algorithm 3 = SLM algorithm

tonsil_wnn <- FindClusters(tonsil_wnn, resolution = c(0.05,0.075),algorithm = 3, graph.name = "wsnn",verbose = FALSE)
print(colnames(tonsil_wnn@meta.data))
```

```{r fig.wide=TRUE}
vars <- str_subset(colnames(tonsil_wnn@meta.data), "^wsnn_res")
clusters_gg <- purrr::map(vars, function(x) {
  p <- DimPlot(
    tonsil_wnn,
    group.by = x,
    reduction = "wnn.umap",
    pt.size = 0.1, label = T
  )
  p 
})
clusters_gg
```


"Naive/Memory B-cell", "CD4+ T-cell","DZ/GC","LZ/GC","CD8+ T-cell","PC","Monocytes","CD4+ T-cell"

## Find all Markers 

```{r}
Idents(tonsil_wnn)<-"wsnn_res.0.075"
tonsil_markers_075<-FindAllMarkers(object = tonsil_wnn, only.pos = TRUE)
write.csv(tonsil_markers_075,file=paste0("~/Documents/multiome_tonsil_Lucia/results/tables/", "tonsil_markers_075_nocluster6n7.csv"))
```

### Get top n for each cluster

**Resolution 0.075**
 
```{r}
tonsil_markers_075 %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC) %>% write.csv(.,file=paste0("~/Documents/multiome_tonsil_Lucia/results/tables/", "top10_tonsil_markers_075_nocluster7.csv.csv"))

tonsil_markers_075 %>% group_by(cluster) %>% top_n(n = 5, wt = avg_log2FC) %>% write.csv(.,file=paste0("~/Documents/multiome_tonsil_Lucia/results/tables/", "top5_tonsil_markers_075_nocluster7.csv.csv"))

top7_tonsil_markers_075<-tonsil_markers_075 %>% group_by(cluster) %>% top_n(n = 7, wt = avg_log2FC)
top5_tonsil_markers_075<-tonsil_markers_075 %>% group_by(cluster) %>% top_n(n = 5, wt = avg_log2FC)
top10_tonsil_markers_075<-tonsil_markers_075 %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)
```

```{r}
df_top5<-as.data.frame(top5_tonsil_markers_075)
kbl(df_top5,caption = "Table of the top 5 marker of each cluster resolution 0.005") %>%
  kable_paper("striped", full_width = F)
```

# Markers exploration

```{r}
markers_gg <- function(x){purrr::map(x, function(x) {
  
  p <- FeaturePlot(
    tonsil_wnn,
    features = x,
    reduction = "wnn.umap",
    pt.size = 0.1
  )
  p
})}
```

**MARKERS**

Immature B cells express CD19, CD 20, CD34, CD38, and CD45R,
T-cell receptor/CD3 complex (TCR/CD3 complex) 


* **T-cells** (identified by high expression of CD3D and CD3E).
* **monocytes** (identified by high expression of LYZ and S100A8).
* **naive B-cells** (identified by high expression of CD79A, CD79B and BLNK).
* **plasma cells** (identified by B-cell and proliferation markers, such as TOP2A or MKI67).
* poor-quality cells (identified by high mitochondrial expression). If a cell has pores in the membrane due to cell lysis, the cytosolic mRNA will leak out of the cell; but if the diameter of mitochondria is higher than the pores, they will get trapped inside the cell.


`


```{r}
monocytes_markers<-c("LYZ","S100A8")

naive_markers<-c("CD79A", "CD79B", "BLNK")

bib_Bcell_markers<-c("CD19","CR2","MS4A1","RALGPS2","CD79A")
bib_Tcell_markers<-c("CD3E","CD4","CD8A","FOXP3","IL17A")

markers_bcell<-c("BANK1","ARHGAP24","ADAM28","MARCH1","RAPGEF5","AFF2","RGS13","LPP","IGHG1","IGLC1","SLC8A1","LYZ","PLXDC2","FAM160A1","IGHA1","IGLC2", "SETBP1","ENTPD1","COL19A1","CCSER1")

markers_tcell<-c("INPP4B","FYB1","LEF1","IL7R","IL6ST","CCL5","GNLY","NKG7","DTHD1","RUNX2", "FOXP3","CD8A","IL17A","CD2")

```

## Bibliography  markers

### T-cells

CD8+ T cell markers:"CD3D", "CD8A"
NK cell markers:"GNLY", "NKG7"

```{r}
markers_gg (bib_Tcell_markers)
```

### B cells


```{r}
markers_gg(bib_Bcell_markers)
```


#### Monocytes 

```{r}
markers_gg(monocytes_markers)
```

#### Naive 

```{r}
markers_gg (naive_markers)

```

```{r}
markers_gg (naive_markers)

```

### Iñaki Markers 

```{r}
m<-c("PRDM1","XBP1","IRF4","MEF2B","BCL6")

DZ<-c("SUGCT", "CXCR4", "AICDA")

LZ<- c("CD83","BCL2A1")

GC<- c("MEF2B", "BCL6","IRF4")

PC<- c("PRDM1","SLAMF7", "MZB1", "FKBP11")


markers_gg(m)

```



#### Dark Zone

```{r}
markers_gg(DZ)

```

#### Light Zone

```{r}
markers_gg(LZ)

```

#### Germinal Center


```{r}
markers_gg(GC)

```

#### Plasma cell


```{r}
markers_gg(PC)

```

### Pauli markers


```{r}
naive_mem_bcell<-c("BANK1", "FCER2")
cd4_tcell<-c("CD3D", "IL7R")
dz_gc_bcell<-c("MKI67", "TOP2A")
lz_gc_bcell<-c("MARCKSL1", "RGS13", "LMO2", "CCDC88A")	
cytotoxic<-c("GNLY", "NKG7", "GZMK", "CD8A")
memory_bcell<-c(	"FCRL4", "FCRL5", "PLAC8", "SOX5")
pc<-c("IGHG1", "IGHA1", "JCHAIN", "XBP1")
myeloid<-c("LYZ", "S100A8")	
poor_q_doublets <-c("FDCSP", "CLU", "CXCL13", "CR2")
doublet_proliferative_tcell<-c("MT2A", "CD3D", "TRAC", "PCNA")
Unk<-c("PTPRCAP", "CD37", "CD74")	
PDC<-c("PTCRA", "LILRA4", "IRF7")
```

```{r}
markers_gg(naive_mem_bcell)
```
```{r}
markers_gg(cd4_tcell)
```
```{r}

markers_gg(dz_gc_bcell)
```

```{r}
markers_gg(lz_gc_bcell)
```
```{r}
markers_gg(cytotoxic)
```
Cytotoxic T cells are effector cells that destroy virus-infected cells, tumor cells, and tissue grafts that exist in the cytosol, or contiguous nuclear compartment. The cells are also known as CD8+ T 

```{r}

markers_gg(pc)
```
```{r}
markers_gg(myeloid)
```

```{r}
markers_gg(poor_q_doublets)
```
```{r}
markers_gg(doublet_proliferative_tcell)
```
```{r}

markers_gg(Unk)
```
```{r}

markers_gg(PDC)

```

# Rename clusters


```{r}
cell.num <- table(Idents(tonsil_wnn_without_doublet))
cell.num
```

```{r}
new.cluster.ids <- c("Naive/Memory B-cell", "CD4+ T-cell","DZ/GC","LZ/GC","CD8+ T-cell","PC","Monocytes","CD4+ T-cell")
names(new.cluster.ids) <- levels(tonsil_wnn)
tonsil_wnn <- RenameIdents(tonsil_wnn, new.cluster.ids)
DimPlot(tonsil_wnn, reduction = "wnn.umap", label = TRUE, pt.size = 0.5) 
```




## Mithocondrial


```{r}
FeaturePlot(
    tonsil_wnn,
    features = "percent.mt",
    reduction = "wnn.umap",
    pt.size = 0.1
  )
```

```{r}
VlnPlot(tonsil_wnn, features = "percent.mt", group.by = "wsnn_res.0.05", pt.size=0)
```
```{r}
VlnPlot(tonsil_wnn, features = "percent_ribo", group.by = "wsnn_res.0.05", pt.size=0)
```

## Doublet score


```{r}
FeaturePlot(
    tonsil_wnn,
    features = "doublet_scores",
    reduction = "wnn.umap",
    pt.size = 0.1
  )
```
```{r eval=FALSE, include=FALSE}
tonsil_wnn<-FindSubCluster(
  tonsil_wnn,
  5,
  graph.name="wsnn",
  subcluster.name = ,"sub.cluster5",
  resolution = 0.25,
  algorithm = 3
)

```
```{r}
vars <- str_subset(colnames(tonsil_wnn@meta.data), "^sub.cluster")
clusters_gg <- purrr::map(vars, function(x) {
  p <- DimPlot(
    tonsil_wnn,
    group.by = x,
    reduction = "wnn.umap",
    pt.size = 0.1, label = T
  )
  p
})
clusters_gg 
```

#Link genes

```{r}
DefaultAssay(tonsil_wnn) <- "peaks"

# first compute the GC content for each peak
tonsil_wnn <- RegionStats(tonsil_wnn, genome = BSgenome.Hsapiens.UCSC.hg38)

# link peaks to genes
tonsil_wnn <- LinkPeaks(
  object = tonsil_wnn,
  peak.assay = "peaks",
  expression.assay = "RNA"
)
```
# Rename clusters

```{r}
new.cluster.ids <- c("B memory/naive","CD8 T","DZ/GC","LZ/GC","CD4 T","PC","monocytes","T cell pro")
names(new.cluster.ids) <- levels(tonsil_wnn)
tonsil_wnn <- RenameIdents(tonsil_wnn, new.cluster.ids)
DimPlot(tonsil_wnn, reduction = "wnn.umap", label = TRUE, pt.size = 0.5) 
```

"B memory/naive","CD8 T",DZ/GC","LZ/GC","CD4 T","PC","monocytes"
```{r}
idents.plot <- c("B memory/naive","CD8 T","DZ/GC","LZ/GC","CD4 T","PC","monocytes","T cell pro")

m<-c("PRDM1","XBP1","IRF4","MEF2B","BCL6", "SUGCT", "CXCR4","CD83","BCL2A1","MEF2B", "BCL6","IRF4","PRDM1","SLAMF7")

coverage_gg <- function(x){purrr::map(x, function(x) {
  
  p <- CoveragePlot(
  object = tonsil_wnn,
  region = x,
  features = x,
  expression.assay = "RNA",
  idents = idents.plot,
  extend.upstream = 5000,
  extend.downstream = 10000
  )
  p
})}
coverage_gg(m)
```


```{r}

p1 <- CoveragePlot(
  object = tonsil_wnn,
  region = "ANK3",
  features = "ANK3",
  expression.assay = "RNA",
  idents = idents.plot,
  extend.upstream = 8000,
  extend.downstream = 1000000
)
  
p2 <- CoveragePlot(
  object = tonsil_wnn,
  region = "PRDM1",
  features = "PRDM1",
  expression.assay = "RNA",
  idents = idents.plot,
  extend.upstream = 8000,
  extend.downstream = 10000
)
p3 <- CoveragePlot(
  object = tonsil_wnn,
  region = "MEF2B",
  features = "MEF2B",
  expression.assay = "RNA",
  idents = idents.plot,
  extend.upstream = 8000,
  extend.downstream = 10000
)
p1
p2
p3
#patchwork::wrap_plots(p1, p2, ncol = 1)
```

# Save

```{r}
saveRDS(tonsil_wnn, path_to_save)
```

# Session Information

```{r}
sessionInfo()
```
