---
title: "6-Analysing clusters without Doublets"
author: 
- name: "Maria Lucia Romero Rivero"
  email: MLROMEROR@clinic.cat
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
    fig_captions: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pre-processing 

## Load packages


```{r message=FALSE, warning=FALSE}
library("RColorBrewer")
library(Signac)
library(Seurat)
library(GenomicRanges)
library(future)
#library(SeuratWrappers)
library(harmony)
library(EnsDb.Hsapiens.v86)
library(stringr)
library(dplyr)
library(ggplot2)
library(patchwork)
library(kableExtra)
library(tidyverse)

set.seed(173)
```


## Parameters

```{r}
# Paths
path_to_obj <- ("~/Documents/multiome_tonsil_Lucia/results/R_objects/12.tonsil_multiome_without_cluster6n7_doublets_normalized.rds")

path_to_markers<-here::here("~/Documents/multiome_tonsil_Lucia/results/tables/tonsil_markers_075_nocluster6n7.csv")
# Thresholds
```


## Load data

```{r}
tonsil_wnn_without_cluster7 <- readRDS(path_to_obj)

tonsil_markers_05<-read_csv(path_to_markers)
```

### Get top n markers of each cluster


**Resolution 0.05**
 
```{r}
tonsil_markers_05 %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC) %>% write.csv(.,file=paste0("~/Documents/multiome_tonsil_Lucia/results/tables/", "top10_tonsil_markers_no_doublets_05.csv"))

tonsil_markers_05 %>% group_by(cluster) %>% top_n(n = 5, wt = avg_log2FC) %>% write.csv(.,file=paste0("~/Documents/multiome_tonsil_Lucia/results/tables/", "top5_tonsil_markers_no_doublets_05.csv"))


top5_tonsil_markers_05<-tonsil_markers_05 %>% group_by(cluster) %>% top_n(n = 5, wt = avg_log2FC)
top7_tonsil_markers_05<-tonsil_markers_05 %>% group_by(cluster) %>% top_n(n = 7, wt = avg_log2FC)

top10_tonsil_markers_05<-tonsil_markers_05 %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)
```

```{r}
df_top5<-as.data.frame(top5_tonsil_markers_05)
kbl(df_top5,caption = "Table of the top 5 marker of each cluster resolution 0.005") %>%
  kable_paper("striped", full_width = F)
```

```{r}
df_top7<-as.data.frame(top7_tonsil_markers_05)
df_mark<-as.data.frame(tonsil_markers_05)
kbl(df_top7,caption = "Table of the top 5 marker of each cluster resolution 0.005") %>%
  kable_paper("striped", full_width = F)
```


```{r}
#install.packages("htmlwidgets", type = "binary")
#install.packages("DT", type = "binary")

DT::datatable(df_top7)

```
```{r}
DT::datatable(df_mark)

```
```{r}
markerGenes <- unique(tonsil_markers_05$gene)

geneSym <- ifelse(test = !grepl('NA', markerGenes), 
             yes = sub('.*?-', '', markerGenes),
             no = sub('-.*', '', markerGenes))


dot.10 <- DotPlot(tonsil_wnn_without_cluster7, features = unique(top10_tonsil_markers_05$gene),cols = 'RdBu', cluster.idents = T) + theme(axis.text.x = element_text( size = 10, vjust = 0.8, hjust = 0.8)) + scale_x_discrete(labels= geneSym)+ggtitle("res 0.05 top 10 of each cluster")

dot.5 <- DotPlot(tonsil_wnn_without_cluster7, features = unique(top5_tonsil_markers_05$gene),cols = 'RdBu', cluster.idents = T) + theme(axis.text.x = element_text( size = 10, vjust = 0.8, hjust = 0.8)) + scale_x_discrete(labels= geneSym)+ggtitle("res 0.05 top 5 of each cluster")

```

```{r}
dot.10 +
  coord_flip() +
  theme(axis.title = element_blank(), axis.text.y = element_text(size = 5))
dot.5 +
  coord_flip() +
  theme(axis.title = element_blank(), axis.text.y = element_text(size = 7))
```
```{r}
top7mark_cluster0<-top7_tonsil_markers_05[["gene"]][1:7]
top7mark_cluster1<-top7_tonsil_markers_05[["gene"]][8:14]
top7mark_cluster2<-top7_tonsil_markers_05[["gene"]][15:21]
top7mark_cluster3<-top7_tonsil_markers_05[["gene"]][22:28]
top7mark_cluster4<-top7_tonsil_markers_05[["gene"]][29:35]
top7mark_cluster5<-top7_tonsil_markers_05[["gene"]][36:42]
top7mark_cluster6<-top7_tonsil_markers_05[["gene"]][43:49]
top7mark_cluster7<-top7_tonsil_markers_05[["gene"]][50:56]

```



# Markers exploration
```{r}
markers_gg <- function(x){purrr::map(x, function(x) {
  
  p <- FeaturePlot(
    tonsil_wnn_without_cluster7,
    features = x,
    reduction = "wnn.umap",
    pt.size = 0.1
  )
  p
})}
```

```{r}
m<-c("PRDM1","XBP1","IRF4","MEF2B","BCL6")

DZ<-c("SUGCT", "CXCR4", "AICDA")

LZ<- c("CD83","BCL2A1")

GC<- c("MEF2B", "BCL6","IRF4")

PC<- c("PRDM1","SLAMF7", "MZB1", "FKBP11")

```


## cluster 0; memory and naive b cells 



```{r}
markers_gg(top7mark_cluster0)
```
```{r}
naive_markers<-c("CD79A", "CD79B", "BLNK")
memory_markers<-c("CD27")

```

```{r}
markers_gg(naive_markers)
```

```{r}
markers_gg(memory_markers)

```

```{r}
markers_gg(c("MS4A1","NT5E"))
```

## cluster 1: naive CD4 T-celL

IL6ST: naive CD4 T-cel

CCR7, CD62L, and CD45RA
```{r}
cd4<- c("CCR7")
```

```{r}
markers_gg(top7mark_cluster1)
```
```{r}
markers_gg(cd4)
```

## cluster 2: GC, DZ


```{r}
markers_gg(top7mark_cluster2)
```

```{r}
markers_gg(GC)

```

```{r}
markers_gg(DZ)

```

## cluster 3: GC, LZ

```{r}
markers_gg(top7mark_cluster3)
```
```{r}
markers_gg(LZ)
```

## cluster 4: NK

ccl5:cd8t cell, nk

AOAH: NK
GNLY:NK


```{r}
markers_gg(top7mark_cluster4)
```

```{r}
markers_gg("KIR2DL4")
```

## cluster 5: PC


```{r}
markers_gg(top7mark_cluster5)
```

```{r}
markers_gg( "MYO1E")
```

```{r}
markers_gg(PC)
```


## cluster 6: monocites

```{r}
markers_gg(top7mark_cluster6)
```

```{r}
monocytes_markers<-c("LYZ","S100A8")
```

```{r}
markers_gg(monocytes_markers)
```

## cluster 7

```{r}
markers_gg(top7mark_cluster7)
```

# Number of cell in each cluster


```{r}
DimPlot(tonsil_wnn_without_cluster7, reduction = "wnn.umap", label = TRUE, pt.size = 0.5) 

```



```{r}
cell.num <- table(Idents(tonsil_wnn_without_cluster7))
cell.num
```

# Rename clusters


```{r}
cell.num <- table(Idents(tonsil_wnn_without_cluster7))
cell.num
```

```{r}
new.cluster.ids <- c("Naive/MBC", "Naive CD4 T-celL","GC/DZ", "GC/LZ", "NK T-cell", "PC", "Monocytes","NI")
names(new.cluster.ids) <- levels(tonsil_wnn_without_cluster7)
tonsil_wnn_without_cluster7 <- RenameIdents(tonsil_wnn_without_cluster7, new.cluster.ids)
DimPlot(tonsil_wnn_without_cluster7, reduction = "wnn.umap", label = TRUE, pt.size = 0.5) 
```



# Bibliography  markers

**MARKERS**

Immature B cells express CD19, CD 20, CD34, CD38, and CD45R,
T-cell receptor/CD3 complex (TCR/CD3 complex) 


* **T-cells** (identified by high expression of CD3D and CD3E).
* **monocytes** (identified by high expression of LYZ and S100A8).
* **naive B-cells** (identified by high expression of CD79A, CD79B and BLNK).
* **plasma cells** (identified by B-cell and proliferation markers, such as TOP2A or MKI67).
* poor-quality cells (identified by high mitochondrial expression). If a cell has pores in the membrane due to cell lysis, the cytosolic mRNA will leak out of the cell; but if the diameter of mitochondria is higher than the pores, they will get trapped inside the cell.

DZ: SUGCT, CXCR4, AICDA

LZ: CD83, BCL2A1

GC total: MEF2B, BCL6, IRF4

PC: PRDM1, SLAMF7, MZB1, FKBP11


```{r}
canonical_bcell_markers <-c("CD34", "CD38", "CD19")

monocytes_markers<-c("LYZ","S100A8")

naive_markers<-c("CD79A", "CD79B", "BLNK")

bib_Bcell_markers<-c("CD19","CR2","MS4A1","RALGPS2","CD79A")
bib_Tcell_markers<-c("CD3E","CD4","CD8A","FOXP3","IL17A")

markers_bcell<-c("BANK1","ARHGAP24","ADAM28","MARCH1","RAPGEF5","AFF2","RGS13","LPP","IGHG1","IGLC1","SLC8A1","LYZ","PLXDC2","FAM160A1","IGHA1","IGLC2", "SETBP1","ENTPD1","COL19A1","CCSER1")

markers_tcell<-c("INPP4B","FYB1","LEF1","IL7R","IL6ST","CCL5","GNLY","NKG7","DTHD1","RUNX2", "FOXP3","CD8A","IL17A","CD2")

```



## B cells

```{r}
markers_gg(naive_markers)
```


```{r}
markers_gg(bib_Bcell_markers)
 
```


## T-cells

CD8+ T cell markers:"CD3D", "CD8A"
NK cell markers:"GNLY", "NKG7"

```{r}
markers_gg(bib_Tcell_markers)
  
```


## Pauli markers


```{r}
naive_mem_bcell<-c("BANK1", "FCER2")
cd4_tcell<-c("CD3D", "IL7R")
cytotoxic<-c("GNLY", "NKG7", "GZMK", "CD8A")
myeloid<-c("LYZ", "S100A8")	

dz_gc_bcell<-c("MKI67", "TOP2A")
lz_gc_bcell<-c("MARCKSL1", "RGS13", "LMO2", "CCDC88A")	
memory_bcell<-c(	"FCRL4", "FCRL5", "PLAC8", "SOX5")
pc<-c("IGHG1", "IGHA1", "JCHAIN", "XBP1")
poor_q_doublets <-c("FDCSP", "CLU", "CXCL13", "CR2")
doublet_proliferative_tcell<-c("MT2A", "CD3D", "TRAC", "PCNA")
Unk<-c("PTPRCAP", "CD37", "CD74")	
PDC<-c("PTCRA", "LILRA4", "IRF7")
```

```{r}
markers_gg(naive_mem_bcell)
```

```{r}
markers_gg(cd4_tcell)
```

```{r}

markers_gg(dz_gc_bcell)
```

```{r}
markers_gg(lz_gc_bcell)
```
```{r}
markers_gg(cytotoxic)
```

Cytotoxic T cells are effector cells that destroy virus-infected cells, tumor cells, and tissue grafts that exist in the cytosol, or contiguous nuclear compartment. The cells are also known as CD8+ T 

```{r}
markers_gg(pc)
```

```{r}
markers_gg(myeloid)
```

```{r}
markers_gg(poor_q_doublets)
```

```{r}
markers_gg(doublet_proliferative_tcell)
```

 Proliferating cell nuclear antigen (PCNA)
 
```{r}

markers_gg(Unk)
```
```{r}

markers_gg(PDC)

```

```{r}

markers_gg("FOXO1")

```

"NFKB1": CLUSTER 2


```{r}
markers_gg( "MYO1E")
```


## Doublet score


```{r}
FeaturePlot(
    tonsil_wnn_without_cluster7,
    features = "doublet_scores",
    reduction = "wnn.umap",
    pt.size = 0.1
  )
```


# CellCycleScoring

```{r}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

tonsil_wnn_without_cluster7 <- CellCycleScoring(tonsil_wnn_without_cluster7, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

head(tonsil_wnn_without_cluster7[[]])

```

```{r}
print(x = tonsil_wnn_without_cluster7[["pca"]], 
      dims = 1:10, 
      nfeatures = 5)
```
PCNA: Proliferating cell nuclear antigen

```{r}
# Visualize the distribution of cell cycle markers across
RidgePlot(tonsil_wnn_without_cluster7, features = c("PCNA", "TOP2A", "MCM6", "MKI67"), ncol = 2)
```
```{r}
tonsil_wnn_without_cluster7 <- RunPCA(tonsil_wnn_without_cluster7, features = c(s.genes, g2m.genes))
tonsil_wnn_without_cluster7 <- RunUMAP(object = tonsil_wnn_without_cluster7,
  nn.name = "weighted.nn",
  reduction.name = "wnn.umap",
  reduction.key = "wnnUMAP_" )
```

```{r}
DimPlot(tonsil_wnn_without_cluster7,
    reduction = "wnn.umap",
    pt.size = 0.1, label = T, split.by = "age_group")
```

```{r}
DimPlot(tonsil_wnn_without_cluster7,
    reduction = "wnn.umap",
    pt.size = 0.1, label = T)
```





