---
title: "Filtering MNB_NBC cluster"
author: 
- name: "Maria Lucia Romero Rivero"
  email: MLROMEROR@clinic.cat
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
    fig_captions: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pre-processing 

## Load package


```{r message=FALSE, warning=FALSE}
library("RColorBrewer")
library(Signac)
library(Seurat)
library(GenomicRanges)
library(future)
library(harmony)
library(EnsDb.Hsapiens.v86)
library(stringr)
library(dplyr)
library(ggplot2)
library(patchwork)
library(kableExtra)
library(tidyverse)
library(EnsDb.Hsapiens.v86)
library(BSgenome.Hsapiens.UCSC.hg38)
set.seed(123)
```

# Parameters

```{r}
# Paths
path_to_obj <- ("~/Documents/multiome_tonsil_Lucia/results/R_objects/15.1.tonsil_multiome_bcells_Linkpeaks_df_normalized.rds")
path_to_save<-("~/Documents/multiome_tonsil_Lucia/results/R_objects/18.tonsil_multiome_bcell_level_1.rds")
```


## Load data

```{r}
tonsil_wnn_bcell <- readRDS(path_to_obj)
```

# Rename

```{r}
Idents(tonsil_wnn_bcell)<-"wsnn_res.0.075"
new.cluster.ids <- c("MBC","NBC", "GC/DZ", "GC/LZ","PC","MBC_NBC")

names(new.cluster.ids) <- levels(tonsil_wnn_bcell)
tonsil_wnn_bcell <- RenameIdents(tonsil_wnn_bcell, new.cluster.ids)
tonsil_wnn_bcell[["annotation_level_1"]] <- Idents(object = tonsil_wnn_bcell)
DimPlot(tonsil_wnn_bcell, reduction = "wnn.umap", pt.size = 0.5,cols = c("#a6cee3", "#1f78b4","#b2df8a", 
             "#33a02c", "#fb9a99","#e31a1c")) + ggtitle("Joint UMAP")
```


# tonsil_wnn_bcell set of MBC, NBC and MBC_NBC

```{r echo=TRUE}
tonsil_wnn_bcell_filter<-tonsil_wnn_bcellset(x = tonsil_wnn_bcell, idents = "MBC_NBC", invert = TRUE)
```

```{r}
DimPlot(tonsil_wnn_bcell_filter, reduction = "wnn.umap", pt.size = 0.5,cols = c("#a6cee3", "#1f78b4","#b2df8a", 
             "#33a02c", "#fb9a99","#e31a1c")) + ggtitle("Joint UMAP")
```


# Normalzation

## scATAC

### Normalization and linear dimensional reduction

 We exclude the first dimension as this is typically correlated with sequencing depth
Cells cluster completely separately in ATAC without harmony; so run harmony after SVD

RunSVD LSI
```{r}
DefaultAssay(tonsil_wnn_bcell_filter) <- "ATAC"
tonsil_wnn_bcell_filter <- RunTFIDF(tonsil_wnn_bcell_filter)
tonsil_wnn_bcell_filter <- FindTopFeatures(tonsil_wnn_bcell_filter, min.cutoff = "q0")
tonsil_wnn_bcell_filter <- RunSVD(tonsil_wnn_bcell_filter)
```



### Plot the Depth correlation plot

Compute the correlation between total counts and each reduced dimension component.

LSI component is typically highly correlated with sequencing depth. 
The first LSI component often captures sequencing depth (technical variation) rather 
than biological variation. If this is the case, the component should be removed from 
downstream analysis. We can assess the correlation between each LSI component and 
sequencing depth using the DepthCor() function:

For scRNA-seq data we don't typically observe such a strong relationship between the first PC and sequencing depth, and so usually retain the first PC in downstream analyses.


```{r fig.wide=TRUE}

DepthCor(tonsil_wnn_bcell_filter)

```

Here we see there is a very strong correlation between the first LSI component and the total 
number of counts for the cell, so we will perform downstream steps without this component.


### UMAP representation

* dimensional reduction key, specifies the string before the number for the dimension names. UMAP by default
* reduction.name: Name to store dimensional reduction under in the Seurat object


```{r fig.wide=TRUE}
tonsil_wnn_bcell_filter <- RunUMAP(
  tonsil_wnn_bcell_filter,
  dims = 2:40,
  reduction = "lsi",
  reduction.name = "umap.atac",
  reduction.key = "atacUMAP_"
)
atac.umap<-DimPlot(
  tonsil_wnn_bcell_filter,
  reduction = "umap.atac",
  group.by = "library_name",
  pt.size = 0.1
) + ggtitle('scATAC UMAP') + NoLegend()

atac.umap

#split_by: library ,edad, genero
```



## scRNA

### Normalization and linear dimensional reduction- 

### NormalizeData

```{r message=FALSE, warning=FALSE}
DefaultAssay(tonsil_wnn_bcell_filter) <- "RNA"
tonsil_wnn_bcell_filter <- NormalizeData(
  tonsil_wnn_bcell_filter,
  normalization.method = "LogNormalize",
  scale.factor = 1e4
)

tonsil_wnn_bcell_filter <- tonsil_wnn_bcell_filter %>%
  FindVariableFeatures(nfeatures = 3000) %>%
  ScaleData() %>% 
  RunPCA() 
```



```{r}
PCAPlot(tonsil_wnn_bcell_filter,
  group.by = "library_name")

```


```{r}
ElbowPlot(object = tonsil_wnn_bcell_filter)
```


### UMAP representation

```{r}
tonsil_wnn_bcell_filter <- RunUMAP(
  tonsil_wnn_bcell_filter,
  dims = 1:30,
  reduction = "pca",
  reduction.name = "umap.rna",
  reduction.key = "rnaUMAP_"
)
rna.umap<-DimPlot(
  tonsil_wnn_bcell_filter,
  reduction = "umap.rna",
  group.by = "library_name",
  pt.size = 0.1) + NoLegend() + ggtitle('scRNA UMAP')

rna.umap
```


```{r}
atac.umap + rna.umap
```


hacer primero harmony, quitar batch effect. 
atac y rna. harmony

# Run Harmony 

Pass the Seurat object to the RunHarmony function and specify which variable to integrate 
out. A Seurat object is returned with corrected Harmony coordinates.

## scATAC

```{r}
DefaultAssay(tonsil_wnn_bcell_filter) <- "ATAC"
tonsil_wnn_bcell_filter <- RunHarmony(
  object = tonsil_wnn_bcell_filter,
  reduction = "lsi",
  dims = 2:40,
  group.by.vars = "library_name",
  assay.use = "ATAC",
  project.dim = FALSE,
  reduction.save = "harmony_ATAC"
)
```


### UMAP representation

```{r}
tonsil_wnn_bcell_filter <- RunUMAP(
  tonsil_wnn_bcell_filter,
  dims = 2:40,
  reduction = "harmony_ATAC",
  reduction.name = "umap.atac",
  reduction.key = "atacUMAP_"
)

Harm_peak<-DimPlot(
  tonsil_wnn_bcell_filter,
  reduction = "umap.atac",
  group.by = "library_name",
  pt.size = 0.1
) + NoLegend() + ggtitle('ATAC Harmony')
```

# scRNA

```{r message=FALSE, warning=FALSE, include=FALSE}
DefaultAssay(tonsil_wnn_bcell_filter) <- "RNA"
tonsil_wnn_bcell_filter <- RunHarmony(
  object = tonsil_wnn_bcell_filter,
  reduction = "pca",
  dims = 1:30,
  group.by.vars = "library_name",
  assay.use = "RNA",
  project.dim = FALSE,
  reduction.save = "harmony_rna"
)
```

```{r}
tonsil_wnn_bcell_filter <- RunUMAP(
  tonsil_wnn_bcell_filter,
  dims = 2:40,
  reduction = "harmony_rna",
  reduction.name = "umap.rna",
  reduction.key = "rnaUMAP_"
)
Harm_rna<-DimPlot(
  tonsil_wnn_bcell_filter,
  reduction = "umap.rna",
  group.by = "library_name",
  pt.size = 0.1
) + NoLegend() + ggtitle('RNA Harmony')
```

## scATAC and RNAseq Harmony

```{r}
Harm_peak+Harm_rna 
```


## Joint UMAP visualization

**FindNeighbors**

Constructs a Shared Nearest Neighbor (SNN) Graph for a given dataset. We first 
determine the k-nearest neighbors of each cell. We use this knn graph to construct 
the SNN graph by calculating the neighborhood overlap (Jaccard index) between 
every cell and its k.param nearest neighbors.

```{r}
# build a joint neighbor graph using both assays
tonsil_wnn_bcell_filter <- FindMultiModalNeighbors(
  object = tonsil_wnn_bcell_filter,
  reduction.list = list("harmony_ATAC", "harmony_rna"),
  dims.list = list(2:40, 1:30), modality.weight.name = "Joint_snn_umap"
  )

# build a joint UMAP visualization

tonsil_wnn_bcell_filter <- RunUMAP(
  object = tonsil_wnn_bcell_filter,
  nn.name = "weighted.nn",
  reduction.name = "wnn.umap",
  reduction.key = "wnnUMAP_")
```

```{r}
joint.umap<- DimPlot(tonsil_wnn_bcell_filter, label = FALSE, group.by = "library_name", pt.size = 0.1,  reduction = "wnn.umap") + plot_annotation(title = 'Joint UMAP')+ ggtitle('Joint UMAP by library name') + NoLegend()

joint.umap
```


```{r}
joint.umap_age<- DimPlot(tonsil_wnn_bcell_filter, label = FALSE, split.by = "age_group", pt.size = 0.1,  reduction = "wnn.umap") + plot_annotation(title = 'Joint UMAP ')+ ggtitle('Joint UMAP age ') 

joint.umap_hospital<- DimPlot(tonsil_wnn_bcell_filter, label = FALSE, split.by = "hospital", pt.size = 0.1,  reduction = "wnn.umap") + plot_annotation(title = 'Joint UMAP ')+ ggtitle('Joint UMAP age ') 

joint.umap_age
joint.umap_hospital
```


## FindClusters

```{r}
#find cluster algorithm 3 = SLM algorithm

tonsil_wnn_bcell_filter <- FindClusters(tonsil_wnn_bcell_filter, resolution = c(0.025,0.03,0.05,0.075),algorithm = 3, graph.name = "wsnn",verbose = FALSE)
print(colnames(tonsil_wnn_bcell_filter@meta.data))
```

```{r fig.wide=TRUE}
vars <- str_subset(colnames(tonsil_wnn_bcell_filter@meta.data), "^wsnn_res")

clusters_gg <- purrr::map(vars, function(x) {
  p <- DimPlot(
    tonsil_wnn_bcell_filter,
    group.by = x,
    reduction = "wnn.umap",
    pt.size = 0.1,cols = c("#a6cee3", "#1f78b4","#b2df8a", 
             "#33a02c", "#fb9a99")
  )
  p 
})
clusters_gg
```

# Rename

```{r}
Idents(tonsil_wnn_bcell_filter)<-"wsnn_res.0.05"
new.cluster.ids <- c("MBC","NBC", "GC/DZ", "GC/LZ","PC")

names(new.cluster.ids) <- levels(tonsil_wnn_bcell_filter)
tonsil_wnn_bcell_filter <- RenameIdents(tonsil_wnn_bcell_filter, new.cluster.ids)
tonsil_wnn_bcell_filter[["annotation_level_1"]] <- Idents(object = tonsil_wnn_bcell_filter)
DimPlot(tonsil_wnn_bcell_filter, reduction = "wnn.umap", pt.size = 0.5,cols = c("#a6cee3", "#1f78b4","#b2df8a", 
             "#33a02c", "#fb9a99","#e31a1c")) + ggtitle("Joint UMAP")
```

# Peak calling


```{r}
# call peaks using MACS2
DefaultAssay(tonsil_wnn_bcell_filter) <- "ATAC"

peaks <- CallPeaks(tonsil_wnn_bcell_filter, macs2.path = "~/Documents/python3/bin/macs2",group.by="wsnn_res.0.05")

df_bcell <- annoGR2DF(peaks)
write_csv(df_bcell, "~/Documents/multiome_tonsil_Lucia/results/tables/18.df_MACS_annotation_level_1_bcell.csv")
saveRDS(peaks, "~/Documents/multiome_tonsil_Lucia/results/tables/18.MACS_annotation_level_1_bcell.rds")


# remove peaks on nonstandard chromosomes and in genomic blacklist regions
peaks <- keepStandardChromosomes(peaks, pruning.mode = "coarse")
peaks <- subsetByOverlaps(x = peaks, ranges = blacklist_hg38_unified, invert = TRUE)
saveRDS(peaks, "~/Documents/multiome_tonsil_Lucia/results/tables/18.1.MACS_annotation_level_1_bcell_set.rds")

df_bcell <- annoGR2DF(peaks)

write_csv(df_bcell, "~/Documents/multiome_tonsil_Lucia/results/tables/18.1.df_MACS_annotation_level_1_bcell.csv")

# quantify counts in each peak:
## Construct a feature x cell matrix from a genomic fragments file


macs2_counts <- FeatureMatrix(
  fragments = Fragments(tonsil_wnn_bcell_filter),
  features = peaks,
  cells = colnames(tonsil_wnn_bcell_filter)
)

# create a new assay using the MACS2 peak set and add it to the Seurat object

annotation <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)
seqlevelsStyle(annotation) <- "UCSC"

fragments<-Fragments(tonsil_wnn_bcell_filter[["ATAC"]])

tonsil_wnn_bcell_filter[["ATAC"]] <- CreateChromatinAssay(
  counts = macs2_counts,
  fragments = fragments,
  annotation = annotation
)


#Run term frequency inverse document frequency (TF-IDF) normalization on a matrix.
DefaultAssay(tonsil_wnn_bcell_filter) <- "ATAC"
tonsil_wnn_bcell_filter <- RunTFIDF(tonsil_wnn_bcell_filter)
tonsil_wnn_bcell_filter <- FindTopFeatures(tonsil_wnn_bcell_filter, min.cutoff = "q0")
tonsil_wnn_bcell_filter <- RunSVD(tonsil_wnn_bcell_filter)

```


# Doublet

```{r}
FeaturePlot(tonsil_wnn_bcell_filter,reduction = "wnn.umap",features = "doublet_scores", min.cutoff = "q1",
            max.cutoff = "q99")
```

# Find All Markers

```{r}
Idents(tonsil_wnn_bcell_filter)<-"wsnn_res.0.05"
tonsil_markers<-FindAllMarkers(object = tonsil_wnn_bcell_filter, only.pos = TRUE,assay = "RNA")
df_markers<-data.frame(tonsil_markers)
write.csv(tonsil_markers,file=paste0("~/Documents/multiome_tonsil_Lucia/results/tables/", "18_tonsil_markers_level_1.csv"))

```


```{r}
top_7_mnbc_markers<-tonsil_markers %>% group_by(cluster) %>% top_n(n = 7, wt = avg_log2FC)
top10_mnbc_markers<-tonsil_markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)
```

```{r}
df_top7<-as.data.frame(top_7_mnbc_markers)
df_top10<-as.data.frame(top10_mnbc_markers)

df_mark<-as.data.frame(tonsil_markers)
```

## Dynamic table top 7 & 10

```{r}
DT::datatable(df_top7)
```
### Dotplot

```{r}
markers<-c("BANK1", "FCER2","CD3D", "IL7R","MKI67", "TOP2A","MARCKSL1", "RGS13", "LMO2", "CCDC88A","GNLY", "NKG7", "GZMK", "CD8A",	"FCRL4", "FCRL5", "PLAC8", "SOX5","IGHG1", "IGHA1", "JCHAIN", "XBP1","LYZ", "S100A8","MT2A", "CD3D", "TRAC", "PCNA","CD19","CR2","MS4A1","RALGPS2","CD79A")

DefaultAssay(tonsil_wnn_bcell_filter)<-"RNA"

dot <- DotPlot(tonsil_wnn_bcell_filter, features = unique(markers),cols = c("lightgrey", "blue"), cluster.idents = T) + theme(axis.text.x = element_text( size = 10, vjust = 0.8, hjust = 0.8)) + ggtitle("Known markers")

dot +
  coord_flip() +
  theme(axis.title = element_blank(), axis.text.y = element_text(size = 5))


```
```{r}
DoHeatmap(tonsil_wnn_bcell_filter,
          features=markers,
           angle = 0)
```

```{r}
sum(markers %in% rownames(GetAssayData(tonsil_wnn_bcell_filter , slot = 'scale.data')))

```
```{r}
DoHeatmap(tonsil_wnn_bcell_filter, features = top10_mnbc_markers$gene) + NoLegend()
```



```{r}
DT::datatable(df_markers
              )
```

# Save

```{r}
saveRDS(tonsil_wnn_bcell_filter, path_to_save)

```

# Session info

```{r}
sessionInfo()
```