---
title: "1-Normalization"
author: 
- name: "Maria Lucia Romero Rivero"
  email: MLROMEROR@clinic.cat
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
    fig_captions: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Objective

The objective of this notebook is to perform pre-processing and dimensional reduction on both assays independently, using standard approaches for RNA and ATAC-seq data. Then, we will follow the ["Joint RNA and ATAC analysis: 10x multiomic"](https://satijalab.org/signac/articles/tonsil_filtered_multiomic.html) vignette from Signac to obtain a joint visualization using both modalities.

# Pre-processing

## Load packages
```{r}
pkgs <- c(
    "renv",
    "reticulate",
    "png",
    "ggplot2",
    "BiocManager",
    "Seurat"
)

bioc_pkgs <- c(
    "SingleCellExperiment",
    "scater",
    "multtest"
)


# Otherwise do it the normal way
install.packages(pkgs)

# Install Bioconductor packages
BiocManager::install(bioc_pkgs, update = FALSE)
```

```{r message=FALSE, warning=FALSE}
library("RColorBrewer")
library(Signac)
library(Seurat)
library(GenomicRanges)
library(future)
#library(SeuratWrappers)
library(harmony)
library(EnsDb.Hsapiens.v86)
library(stringr)
library(dplyr)
library(ggplot2)
library(patchwork)

set.seed(123)
```


## Parameters

```{r}
path_to_object <- ("/Users/mlromeror/Documents/multiome_tonsil_Lucia/results/R_objects/7.tonsil_filtered_merged_with_metadata.rds")
path_to_save <- ("/Users/mlromeror/Documents/multiome_tonsil_Lucia/results/R_objects/8.tonsil_multiome_integrated_using_wnn.rds")

```



## Load Multiome filtered data with metadata 

```{r}
tonsil_filtered <- readRDS(path_to_object)
tonsil_filtered
```

**change libray names*
```{r}
colnames(tonsil_filtered@meta.data)[16]<-"library_name"
```






## scATAC

### Normalization and linear dimensional reduction

 We exclude the first dimension as this is typically correlated with sequencing depth
Cells cluster completely separately in ATAC without harmony; so run harmony after SVD

RunSVD LSI
```{r}
DefaultAssay(tonsil_filtered) <- "peaks"
tonsil_filtered <- RunTFIDF(tonsil_filtered)
tonsil_filtered <- FindTopFeatures(tonsil_filtered, min.cutoff = "q0")
tonsil_filtered <- RunSVD(tonsil_filtered)
```



### Plot the Depth correlation plot

Compute the correlation between total counts and each reduced dimension component.


LSI component is typically highly correlated with sequencing depth. 
The first LSI component often captures sequencing depth (technical variation) rather 
than biological variation. If this is the case, the component should be removed from 
downstream analysis. We can assess the correlation between each LSI component and 
sequencing depth using the DepthCor() function:

For scRNA-seq data we don't typically observe such a strong relationship between the first PC and sequencing depth, and so usually retain the first PC in downstream analyses.

```{r fig.wide=TRUE}

DepthCor(tonsil_filtered)

```

Here we see there is a very strong correlation between the first LSI component and the total 
number of counts for the cell, so we will perform downstream steps without this component.


### UMAP representation

* dimensional reduction key, specifies the string before the number for the dimension names. UMAP by default
* reduction.name: Name to store dimensional reduction under in the Seurat object


```{r fig.wide=TRUE}
tonsil_filtered <- RunUMAP(
  tonsil_filtered,
  dims = 2:40,
  reduction = "lsi",
  reduction.name = "umap.atac",
  reduction.key = "atacUMAP_"
)
atac.umap<-DimPlot(
  tonsil_filtered,
  reduction = "umap.atac",
  group.by = "library_name",
  pt.size = 0.1
) + ggtitle('scATAC UMAP') + NoLegend()

atac.umap

#split_by: library ,edad, genero
```





## scRNA

### Normalization and linear dimensional reduction- 

### NormalizeData (Log Normalization)

```{r message=FALSE, warning=FALSE}
DefaultAssay(tonsil_filtered) <- "RNA"
tonsil_filtered <- NormalizeData(
  tonsil_filtered,
  normalization.method = "LogNormalize",
  scale.factor = 1e4
)

tonsil_filtered <- tonsil_filtered %>%
  FindVariableFeatures(nfeatures = 3000) %>%
  ScaleData() %>% 
  RunPCA() 
```



```{r}
PCAPlot(tonsil_filtered,
  group.by = "library_name")

```


```{r eval=FALSE, include=FALSE}
PrintPCA(object = tonsil_filtered, pcs.print = 1:5, genes.print = 5, use.full = FALSE)
PCHeatmap(object = tonsil_filtered)
```

```{r}
ElbowPlot(object = tonsil_filtered)
```



### UMAP representation

```{r}
tonsil_filtered <- RunUMAP(
  tonsil_filtered,
  dims = 1:30,
  reduction = "pca",
  reduction.name = "umap.rna",
  reduction.key = "rnaUMAP_"
)
rna.umap<-DimPlot(
  tonsil_filtered,
  reduction = "umap.rna",
  group.by = "library_name",
  pt.size = 0.1) + NoLegend() + ggtitle('scRNA UMAP')

rna.umap
plot(rna.umap)
```


```{r}
atac.umap + rna.umap
```


hacer primero harmony, quitar batch effect. 
atac y rna. harmony

# Harmony Integration (correct data bath effect)

Pass the Seurat object to the RunHarmony function and specify which variable to integrate 
out. A Seurat object is returned with corrected Harmony coordinates.

## scATAC

```{r}
DefaultAssay(tonsil_filtered) <- "peaks"
tonsil_filtered <- RunHarmony(
  object = tonsil_filtered,
  reduction = "lsi",
  dims = 2:40,
  group.by.vars = "library_name",
  assay.use = "peaks",
  project.dim = FALSE,
  reduction.save = "harmony_peaks"
)
```


### UMAP representation

```{r}
tonsil_filtered <- RunUMAP(
  tonsil_filtered,
  dims = 2:40,
  reduction = "harmony_peaks",
  reduction.name = "umap.atac",
  reduction.key = "atacUMAP_"
)

Harm_peak<-DimPlot(
  tonsil_filtered,
  reduction = "umap.atac",
  group.by = "library_name",
  pt.size = 0.1
) + NoLegend() + ggtitle('Peak Harmony')
```

# scRNA

```{r message=FALSE, warning=FALSE, include=FALSE}
DefaultAssay(tonsil_filtered) <- "RNA"
tonsil_filtered <- RunHarmony(
  object = tonsil_filtered,
  reduction = "pca",
  dims = 1:30,
  group.by.vars = "library_name",
  assay.use = "RNA",
  project.dim = FALSE,
  reduction.save = "harmony_rna"
)
```

```{r}
tonsil_filtered <- RunUMAP(
  tonsil_filtered,
  dims = 2:40,
  reduction = "harmony_rna",
  reduction.name = "umap.rna",
  reduction.key = "rnaUMAP_"
)
Harm_rna<-DimPlot(
  tonsil_filtered,
  reduction = "umap.rna",
  group.by = "library_name",
  pt.size = 0.1
) + NoLegend() + ggtitle('RNA Harmony')
```

## scATAC and RNAseq Harmony

```{r}
Harm_peak+Harm_rna + plot_annotation(title = 'Harmony ATAC and RNA UMAP visualization')
```


## Joint UMAP visualization

**FindNeighbors**

Constructs a Shared Nearest Neighbor (SNN) Graph for a given dataset. We first 
determine the k-nearest neighbors of each cell. We use this knn graph to construct 
the SNN graph by calculating the neighborhood overlap (Jaccard index) between 
every cell and its k.param nearest neighbors.

```{r}
# build a joint neighbor graph using both assays
tonsil_filtered <- FindMultiModalNeighbors(
  object = tonsil_filtered,
  reduction.list = list("harmony_peaks", "harmony_rna"),
  dims.list = list(2:40, 1:30), modality.weight.name = "Joint_snn_umap"
  )

# build a joint UMAP visualization

tonsil_filtered <- RunUMAP(
  object = tonsil_filtered,
  nn.name = "weighted.nn",
  reduction.name = "wnn.umap",
  reduction.key = "wnnUMAP_")

joint.umap<- DimPlot(tonsil_filtered, label = FALSE, group.by = "library_name", pt.size = 0.1,  reduction = "wnn.umap") + plot_annotation(title = 'Joint UMAP')+ ggtitle('Joint UMAP by library name') + NoLegend()

joint.umap_age<- DimPlot(tonsil_filtered, label = FALSE, split.by = "age_group", pt.size = 0.1,  reduction = "wnn.umap") + plot_annotation(title = 'Joint UMAP ')+ ggtitle('Joint UMAP age ') 

joint.umap_hospital<- DimPlot(tonsil_filtered, label = FALSE, split.by = "hospital", pt.size = 0.1,  reduction = "wnn.umap") + plot_annotation(title = 'Joint UMAP ')+ ggtitle('Joint UMAP age ') 

```

```{r}
(atac.umap + rna.umap )
```



```{r}
joint.umap 
joint.umap_age
joint.umap_hospital
```





# Save

We will save the resulting object and use it in the following notebook to exclude doublets:

```{r eval=FALSE, include=FALSE}
saveRDS(tonsil_filtered, path_to_save)
```


# Session Information

```{r}
sessionInfo()
```
